---
title: "Hypsithocus no-choice oviposition"
author: "Tom Saunders"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
This notebook contains code used to analyse data from experiments with Hypsithocus hudsonae, an endemic New Zealand pentatomid species, and Trissolcus japonicus, a scelionid parasitoid proposed for release as a biological control agent against brown marmorated stink bug in New Zealand (should the stink bug establish). The underlying dataset is made up of data from experiments conducted by Saunders et al with H. hudsonae and T. japonicus, combined with data from experiments reported by Charles et al (2019) (https://dx.doi.org/10/ggrtkp). I conducted no-choice oviposition experiments under very similar conditions to Charles et al, so that we could eventually pool the data to look at differences in parasitoid emergence between different non-target New Zealand pentatomid species. 

```{r library, include=FALSE}
library(tidyverse)
library(readxl)
library(car)
library(emmeans)
library(lme4)
library(boot)
```

```{r}
# Load data

hyp.df <- read_excel("hypsithocus_raw.xlsx", sheet = "combined_data")
```

```{r}
# Summary of sample size and avg eggs per mass for each pentatomid species x treatment

pents_sum <- hyp.df %>% 
  group_by(species, treatment) %>% 
  summarise(
    egg_masses = n(),
    avg_eggs = mean(n_eggs, na.rm = TRUE),
    )
```

```{r}
# Calculate a range of emergence-related variables, including average percent egg parasitism (the average proportion of eggs within each mass which gave rise to a wasp) and total mass acceptance (the proportion of masses from which at least one wasp emerged)

emerge_avg <- hyp.df %>% 
  group_by(species, treatment) %>% 
  mutate(
    p_pents = n_pents / n_eggs,
    p_wasps = (n_wasps_m + n_wasps_f) / n_eggs,
    p_wasps_f = n_wasps_f / (n_wasps_f + n_wasps_m),
    p_unhatched = (n_eggs - (n_pents + n_wasps_f + n_wasps_m)) / n_eggs,
  ) %>% 
  summarise(
    avg_pents = mean(p_pents),
    avg_wasps = mean(p_wasps),
    avg_ratio_f = mean(p_wasps_f, na.rm = TRUE),
    avg_unhatched = mean(p_unhatched),
    acceptance = (sum(n_wasps_m > 0 | n_wasps_f > 0)) / n(),
    )

# Plot total mass acceptance and average percent egg parasitism

emerge_plot <- emerge_avg %>% 
  filter(treatment %in% c("No-Choice")) %>% 
  pivot_longer(
    cols = c("acceptance", "avg_wasps"),
    names_to = "metric",
    values_to = "proportion",
  )

ggplot(emerge_plot, aes(y = reorder(species, proportion), proportion, fill = metric)) +
  geom_bar(position = "dodge", stat = "identity", show.legend = TRUE) +
  xlab("Proportion") +
  ylab("Species") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.text.y = element_text(face = "italic")) +
  scale_fill_manual(values = c("gray90", "black"), name = "", breaks = c("acceptance", "avg_wasps"), labels = c("Total Mass Acceptance", "Mean Egg Parasitism"))  +
  theme(panel.background = element_rect(fill = "white"),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        plot.background = element_rect(
          fill = "white",
          colour = "black",
          size = 1))


```

```{r}
# Plot spread of percent parasitism for each pentatomid

emergence_spread <- hyp.df %>% 
  filter(treatment %in% c("No-Choice")) %>% 
  mutate(
    emergence = (n_wasps_m + n_wasps_f) / n_eggs,
  ) 

ggplot(emergence_spread, aes(emergence, species)) +
  geom_violin() +
  geom_point()

```

```{r}
# Binomial GLM

# Convert `source`, `species` and `treatment` to factors

str(hyp.df)
hyp.df$source <- as.factor(hyp.df$source)
hyp.df$species <- as.factor(hyp.df$species)
hyp.df$treatment <- as.factor(hyp.df$treatment)
```

```{r}
# Build model: Compare percent egg parasitism between hypsithocus (my data) and other pentatomids tested by Charles et al

# First filter out control obs

hyp2.df <- hyp.df %>% 
  filter(treatment != "Control")

model1 <- glmer(cbind(
  n_wasps_m + n_wasps_f, 
  n_eggs - (n_wasps_m + n_wasps_f)) 
  ~ species + (1|source),  
  data = hyp2.df, 
  family = binomial()
  )

# ANOVA

Anova(model1)
```

```{r}
# Get predicted percent parasitism means and CI to see the magnitude of differences

hyp_means <- emmeans(model1, ~ species) 

# Get post-hoc pairwise comparisons to see where the differences lie

hyp_conmeans <- contrast(hyp_means, method = "pairwise", adjust = "fdr")
write.table(hyp_conmeans, file = "hyp-conmeans.txt", sep = ",", quote = FALSE, row.names = F)

# Plot means and comparisons on the transformed scale

plot(hyp_means)
plot(hyp_conmeans)

```


```{r}
# Back transform the means and CI

hyp_back <- as.data.frame(hyp_means)
means_table <- hyp_back %>%  
  mutate(
    BT_mean=inv.logit(emmean), 
    BT_lower=inv.logit(asymp.LCL), 
    BT_upper=inv.logit(asymp.UCL)) %>%  
  mutate(
    BT_lower_CI = case_when(
      BT_lower == 0 & BT_upper == 1 ~ NA_real_,
      TRUE ~ BT_lower), 
    BT_upper_CI = case_when(
      BT_lower == 0 & BT_upper == 1 ~ NA_real_,
      TRUE ~ BT_upper
))

# Plot back-transformed variables

ggplot(means_table, aes(y = reorder(species, BT_mean), BT_mean)) +
  geom_errorbar(aes(xmin = BT_lower_CI, xmax = BT_upper_CI), width = 0.5) +
geom_point() + xlim(0,1) +
    labs(x = "Probability of parasitoid emergence", y = "Species") +
  theme(axis.text.y = element_text(face = "italic")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.text.y = element_text(face = "italic")) +
  theme(panel.background = element_rect(fill = "white"),
        plot.margin = margin(1, 1, 1, 1, "cm"),
        plot.background = element_rect(
          fill = "white",
          colour = "black",
          size = 1))

```
